Bootstrap learning via modular concept discovery

### 2013-04-20 drafts/ijcai2013/ijcai13.tex
Cleanup and add figures directory.
Changes to ijcai13 tex file.

\maketitle
\begin{abstract}
%\small
 Suppose a learner is faced with a domain of problems about which it
 knows nearly nothing. It does not know the distribution of problems,
 the space of solutions is not smooth, and the reward signal is
 uninformative, providing perhaps a few bits of information but not
 enough to steer it effectively. How can such a learner ever get off
 the ground? A common intuition is that if the solutions to these
 problems share a common structure, and the learner can solve some
 simple problems by brute force, it should be able to extract useful
 components from these solutions and, by composing them, explore the
 solution space more efficiently. Here, we formalize this intuition,
 where the solution space is the space of typed functional programs
 and the gained information is stored as a stochastic grammar over
 programs. We propose an iterative procedure for exploring such
 spaces: in the first step of each iteration, the learner explores a
 finite subset of the domain, guided by a stochastic grammar; in the
 second step, the learner compresses the successful solutions from the
 first step to estimate a new stochastic grammar. We test this
 procedure on symbolic regression and boolean circuit learning and
 show that the learner discovers modular concepts for these
 domains. Whereas the learner is able to solve almost none of the
 posed problems in the procedure's first iteration, it rapidly becomes
 able to solve a large number by gaining abstract knowledge of the
 structure of the solution space.

@@

\section{Introduction}
Many of the problems we want to solve in AI are best cast as search
problems. There are many search algorithms, but they typically exploit
local or topological structure in the search space. For example,
convex optimization and gradient based search methods take
advantage of local smoothness to move in good directions. Local search
algorithms move from one candidate solution to a neighboring
one. Heuristic search techniques take advantage of topological
structure to prune off large parts of the solution space.

However, many of the problems we want solve -- discovering scientific
theories, developing new technologies, figuring out how to bake -- do
not come with such a structured search space. This is particularly
true of problems, such as those mentioned, whose solutions can be
naturally framed as programs. Here we ask how to extract structure
from the space of programs by attempting to solve multiple related
problems in a given domain at once.

We take as inspiration the programmer who -- having reused several
subroutines to solve different but similar simple tasks -- creates a
function library to help manage the complexity of larger projects. The
EC algorithm, which we introduce here, builds a library of reusable
program components and places a distribution over these, effectively
learning from simple tasks a search space structure that enables it
to solve more complex tasks.

@@

Concretely, the EC algorithm works as follows. It begins with a small
library of program primitives and a distribution over programs
synthesized from this library. In each iteration, it generates
programs in order of highest probability according to the library
learned in the previous iteration. From the discovered solutions, it
finds new reusable functions that compress these solutions, and it
re-estimates the distribution over programs by weighting these new
functions according to their frequencies. In using compression as a
guide to the choice of representation, we instantiate the idea that
good representations are those which minimize the description length
of typical elements in the domain.

@@

The remainder of this paper will include two major sections. In the
first, we describe in detail the EC algorithm and design choices we
make in implementing it. We will then present results from two
experiments in the domains of symbolic regression and Boolean
function learning.

@@

We propose the \textbf{EC algorithm}, an iterative procedure to solve this
problem. It will be useful in presenting the algorithm to define the
following terminology: a \textbf{frontier} is the finite set of
expressions that the algorithm considers in any one iteration. The
\textbf{frontier size} $N$ is the prespecified number of elements in
the frontier. We will say that a task is \textbf{hit} by the frontier
if there is a expression in the frontier that solves it.

The EC algorithm maintains a distribution $\mathcal{D}$ over
expressions in $\mathcal{L}$. At each iteration, it
\begin{enumerate}
\item \textbf{explores} the frontier -- enumerating the $N$ most probable expressions
  from the current distribution $\mathcal{D}$ -- and
\item \textbf{compresses} the hit tasks in the frontier to estimate a new
  distribution.
\end{enumerate}

@@

Our best-first policy is to expand the next partial solution
with the highest value. If a partial solution has multiple leaves that
are not primitive combinators, we expand the left most one first,
since it is the choice of left combinator that defines the type of its
right combinator in an application pair. That is, the parent node
applies its constraint on valid solutions via the left child, not the
right child. We want to apply constraints as early as possible, so
expanding left child first prunes more partial solutions.

@@

  \vspace{2mm}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 4 ]
     \end{tikzpicture}
     \caption{(16)}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 5 ]
     \end{tikzpicture}
     \caption{(25)}
  \end{subfigure}
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture} [scale=.6]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 6 ]
     \end{tikzpicture}
     \caption{(36)}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 7 ]
     \end{tikzpicture}
     \caption{(49)}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 8 ]
     \end{tikzpicture}
     \caption{(64)}
  \end{subfigure}
\end{figure}
\captionsetup[subfigure]{labelformat=parens} 
\vspace{-.3cm}
The top row consists of one type of representation of an integer
squared, namely, the integer times itself. The bottom row contains
another type of representation in which the function $f(x)=x^2$,
represented in combinatory logic as $S * I$, is applied to the
integer. The top and bottom row of each expression evaluate to the
same integer. If we consider only the first column, the top tree has 4
unique subtrees and the bottom has 7. But each column adds 3 unique
subtrees to the top row and only 2 to the bottom. So by the fifth
column the top row has more unique subtrees than the bottom row. The
increased compressibility of the bottom row is due to the fact that
the squaring operation has been factored out as an abstraction. This
is an example of how this compressibility metric favors
abstraction when multiple expressions are being considered at once.

@@

Given a set of chosen solution expressions, which we will call the
\emph{solution set}, want to re-estimate our stochastic grammar. Our
inspiration for doing this is the Neville-Manning algorithm for
grammar-based compression~\cite{nevill1997identifying}. The idea of
that compression algorithm is to compress any string by creating a
grammar for that string such that a) no two symbols appear more than
once in the grammar and b) every rule is used more than once. From a
set of expressions we generate a grammar according to these criteria.

@@

\setcounter{subfigure}{0} 
\begin{figure}[ht]
\centering
\includegraphics[width=.8\linewidth]{figures/symreg3frontiers.pdf}
\vspace{-.2cm}
\caption{Learning curves as a function of frontier size.  As frontier
  size is increased, curves plateau closer to $100\%$ performance. A
  baseline search over $150000$ expressions only hits $3\%$ of the
  tasks. \label{fig:symregLearning}}
\vspace{-.3cm}
\end{figure}
\begin{figure}[ht]
\centering
\includegraphics[width=.8\linewidth]{figures/symRegDiffCurricula.pdf}
\caption{How do different task sets affect learning curves? Learning
  curves at frontier size of $10000$ for different task
  sets.} 
\label{fig:symregCurricula}
\vspace{-.5cm}
\end{figure}

@@

In Figure~\ref{fig:symregLearning}, we show performance results as we
vary the frontier size.Changing the frontier size changes the number
of tasks the algorithm identifies correctly over the course of 15
algorithm iterations (all experiments in this work were run for 15
iterations). As a baseline, we enumerated $10000*15=150000$
expressions from the initial grammar. This is the total number of
expressions that a run of the algorithm sees if it has a frontier size
of $10000$ and runs for $15$ iterations. Thus, if the benefit accruing
to the runs with larger frontier sizes is due to simply an increase in
the number of expressions seen, rather than increased learning, we should
see similarly good performance from this baseline run. In fact,
however, the baseline run only hits $27$ of the $1000$ tasks ($3\%$),
whereas our algorithm nears $100\%$ for a frontier of $10000$.

What does the EC algorithm learn in this task?  Inspecting the top
weighted primitives of the final grammars, we find many incrementers
(e.g. (+1), (+2), etc.), several versions of expressions that
translate to functions like $x*f(x)$ and $f(x*f(x))$, and combinations
of these, like $x*(x+3)+3$ and $x*(x+2)$. Informally, we see
expressions that apply functions to themselves, building up complex
functions with relatively few unique primitives.

To what degree is ``conceptual bootstrapping'' responsible for the
improved performance. That is, to what extent does the presence of
simple problems account for the ability to learn complex functions?
One hypothesis is that to succeed on a set of tasks, the set must
contain a ``learnable'' curriculum, a set of tasks that serve as
stepping stones from an impoverished representation to a rich one. We
can test this hypothesis by varying the set of tasks to which the EC
algorithm is exposed. If it is true, then we should see a nonlinear
response to reducing the number of easy tasks, as the curriculum
changes from being learnable to being unlearnable.

@@

\begin{figure}[t]
\includegraphics{./tablealonedoc.pdf}
\vspace{-.5cm}
\caption{Some learned primitives from the circuit based Boolean
  function learning experiment. We highlight compression using $NOT$,
  $E_1$ and $E_2$. \,(a)The three elementary gates used to generate
  the task set. (b) Two higher-order primitives in the induced
  grammar. Note how $E_1$ can be used to define FALSE by taking the
  NOT function as an argument. (c) Additional common Boolean functions
  found in the learned grammar.  \label{table:booleanCircuits}}
\vspace{-.4cm}
\end{figure}

As another demonstration of our approach, we investigate the
EC algorithm's ability to learn Boolean functions using only the logical
primitive NAND and the basic combinators. It is well known that a
boolean circuit can be constructed for any boolean function given only
the NAND gate. Here, the basic combinators take the place of the
wiring normally found in a circuit.

@@

We constructed two task sets. The first of these was constructed
explicitly to contain familiar modular structure. In this set of
tasks, we sampled $1000$ Boolean circuits using AND, OR, and NOT
gates. To accomplish this, we first sampled either 1, 2, 3 inputs,
then between 1 and 5 gates, randomly wiring the inputs of each new
gate to the output of one of the existing gates in the circuit. We
continued this sampling procedure until we had 1000 ``connected''
circuits, i.e., circuits all of whose outputs are wired to an input
except for the last one (the output gate). This resulted in $1000$
tasks consisting of $82$ unique Boolean functions, with a distribution
of tasks as shown in Figure~\ref{fig:booldistr}~(a). In
Figure~\ref{fig:boolLearningCurvesCircuits}, We present learning
curves for frontier sizes 100, 500 and 10000, using the same procedure
as in Section~\ref{sec:symreg}. In Figure~\ref{fig:booldistr}, we show
that the distribution over Boolean functions enumerated from the
grammar over expressions is much more similar to the true function
distribution after learning than before.

\begin{figure}
\vspace{-.3cm}
\includegraphics[width=\linewidth]{figures/circuitDistrComp.pdf}
\vspace{-.6cm}
\caption{Distribution of boolean functions among (a) the set of 1000
  circuits randomly constructed from AND, OR and NOT gates (see text);
  b) the first 1000 expressions from the grammar over expressions
  \emph{before} learning; c) the first 1000 expressions from the
  grammar over expressions \emph{after} learning. Gray levels indicate
  percent of instances (circuits/expressions) which evaluate to that
  function.\label{fig:booldistr}}
\vspace{-.4cm}
\end{figure}

This problem is particularly suitable for inspecting the constituents
of the induced grammar $\mathcal{G}$. We might hope that our algorithm
recovers the constituent logic gates that were used to build up the
tasks. Table~\ref{table:booleanCircuits} shows a few of the top ranked
expressions in the library. These include the basic logic gates; we
also include two higher-order expression $E_1$ and $E_2$, to stress
that the representation the EC algorithm is using allows it to build
concepts that are more expressive that just sub-circuits.

In a second experiment, we include all $272$ Boolean truth tables with
three or fewer inputs as tasks. This is a more difficult
problem. There are two kinds of learning that the EC algorithm might
accomplish: first, many expressions in $\mathcal{L}$ map to a single
Boolean function, so it needs to learn primitives that allow it to
span many different functions instead of generating very simple
functions redundantly. The second kind of learning involves the
distribution of the functions themselves. In the circuit based Boolean
experiment, that structure is apparent in the distribution in
Figure~\ref{fig:booldistr}. In this second experiment, we remove the
second kind of structure. In
Figure~\ref{fig:boolLearningCurvesUniform}, we show 10 runs of the EC
algorithm on the second experiment with a frontier size of $2000$: note how there are several
local minima that the majority of the runs get stuck in with performance
around $50\%$, but several of the runs seem to take different
trajectories to more successful representations.

\vspace{-.3cm}
\begin{figure}[h]
\begin{subfigure}{.1\linewidth}
\caption{\label{fig:boolLearningCurvesCircuits} }
\end{subfigure}
\begin{subfigure}[Before]{.9\linewidth}
\includegraphics[width=.9\linewidth]{figures/booleancircuits3frontiers.pdf}
\end{subfigure}

\vspace{-.32cm}
\begin{subfigure}{.1\linewidth}
\caption{\label{fig:boolLearningCurvesUniform}}
\end{subfigure}
\begin{subfigure}[After]{.9\linewidth}
\includegraphics[width=.9\linewidth]{figures/booleanfunctionsUniform.pdf}

@@

Our specific implementation of this approach involved a number of
choices which should be explored further in future work.  Our claim
here is not that the EC algorithm is optimal, but that it captures key
features of how people solve the problem of learning new systems of
concepts in acquiring deep domain expertise.  Much like a human
researcher investigating a new area of study, our algorithm tries
different models, watches them fail or succeed, and abstracts out the
model elements that capture relevant differences.  It recombines and
reuses successful patterns of reasoning that span many problems in a
domain.

>>>>

\maketitle
\begin{abstract}
%\small
 Suppose a learner is faced with a domain of problems about which it
 knows nearly nothing. It does not know the distribution of problems,
 the space of solutions is not smooth, and the reward signal is
 uninformative, providing perhaps a few bits of information but not
 enough to steer it effectively. How can such a learner ever get off
 the ground? A common intuition is that if the solutions to these
 problems share a common structure, and the learner can solve some
 simple problems by brute force, it should be able to extract useful
 components from these solutions and, by composing them, explore the
 solution space more efficiently. Here, we formalize this intuition,
 where the solution space is the space of typed functional programs
 and the gained information is stored as a stochastic grammar over
 programs. We propose an iterative procedure for exploring such
 spaces: in the first step of each iteration, the learner explores a
 finite subset of the domain, guided by a stochastic grammar; in the
 second step, the learner compresses the successful solutions from the
 first step to estimate a new stochastic grammar. We test this
 procedure on symbolic regression and boolean circuit learning and
 show that the learner discovers modular concepts for these
 domains. Whereas the learner is initially unable to solve almost any
 of the posed problems, it quickly acquires the ability to solve many
 problems by gaining abstract knowledge of the structure of the
 solution space.

@@

\section{Introduction}
A central question for AI is how to create intelligent agents without
building in vast amounts of pre-specified knowledge that is tailored
to one domain or another. Any attempt to solve this problem must
contend with the fact that the less an agent knows about solving
problems in a domain, the worse its attempts are, and the less it
learns about how to solve subsequent problems. For example, imagine
you know nothing about baking but are supplied with a stocked kitchen:
your initial attempts will be uniformly disastrous, and you will learn
little from them, except that you did not create a baked good. A
little bit of knowledge about baking would go a long way towards
making your initial attempts more informative.

Like many of the problems in AI, this one can be cast as a search
problem. There are many search algorithms, but they typically exploit
local or topological structure in the search space. For example,
convex optimization and gradient based search methods take
advantage of local smoothness to move in good directions. Local search
algorithms move from one candidate solution to a neighboring
one. Heuristic search techniques take advantage of topological
structure to prune off large parts of the solution space.

The absense of domain specific knowledge makes exploiting structure
difficult, especially if the structure of the search space is not
smooth. This is particularly true of problems -- such as discovering
scientific theories, developing new technologies, learning how to bake
-- whose solutions can be naturally framed as programs. Here we ask
how to bootstrap domain specific knowledge and extract structure from
the space of programs by attempting to solve multiple related problems
in a given domain at once.

Out approach is inspired by the programmer who -- having reused
several subroutines to solve different but similar simple tasks --
creates a function library to help manage the complexity of larger
projects. The E.C. algorithm, which we introduce here, builds a library
of reusable program components and places a distribution over these,
effectively learning from simple tasks a search space structure that
enables it to solve more complex tasks.

@@

Concretely, the E.C. algorithm works as follows. It begins with a small
library of program primitives and a distribution over programs
synthesized from this library. In each iteration, it generates
programs in order of highest probability according to the library
learned in the previous iteration. From the discovered solutions, it
finds new reusable functions that compress these solutions, and it
re-estimates the distribution over programs by weighting these new
functions according to their frequencies. In using compression as a
guide to the choice of representation, we instantiate the idea that
good representations are those which minimize the description length
of typical elements in the domain.

@@

The remainder of this paper will include two major sections. In the
first, we describe in detail the E.C. algorithm and design choices we
make in implementing it. We will then present results from two
experiments in the domains of symbolic regression and Boolean
function learning.

@@

We propose the \textbf{E.C. algorithm}, an iterative procedure to solve this
problem. It will be useful in presenting the algorithm to define the
following terminology: a \textbf{frontier} is the finite set of
expressions that the algorithm considers in any one iteration. The
\textbf{frontier size} $N$ is the prespecified number of elements in
the frontier. We will say that a task is \textbf{hit} by the frontier
if there is a expression in the frontier that solves it.

The E.C. algorithm maintains a distribution $\mathcal{D}$ over
expressions in $\mathcal{L}$. At each iteration, it
\begin{enumerate}
\item \textbf{explores} the frontier -- enumerating the $N$ most probable expressions
  from the current distribution $\mathcal{D}$ -- and
\item \textbf{compresses} the hit tasks in the frontier to estimate a new
  distribution.
\end{enumerate}

@@

Our best-first policy is to expand the next partial solution with the
highest value. If a partial solution has multiple leaves that are not
primitive combinators, we expand the left most one first, since it is
the choice of left combinator that defines the type of its right
combinator in an application pair. That is, the parent node applies
its constraint on valid solutions via the left child, not the right
child. We want to apply constraints as early as possible, so we expand
the left child first to prune more partial solutions.

@@

  \vspace{2mm}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 4 ]
     \end{tikzpicture}
     \caption{(16)}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 5 ]
     \end{tikzpicture}
     \caption{(25)}
  \end{subfigure}
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture} [scale=.6]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 6 ]
     \end{tikzpicture}
     \caption{(36)}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 7 ]
     \end{tikzpicture}
     \caption{(49)}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.6] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[[  S * ] I ] 8 ]
     \end{tikzpicture}
     \caption{(64)}
  \end{subfigure}
\end{figure}
\captionsetup[subfigure]{labelformat=parens} 
\vspace{-.3cm}

\noindent The top row consists of one type of representation of an integer
squared, namely, the integer times itself. The bottom row contains
another type of representation in which the function $f(x)=x^2$,
represented in combinatory logic as $S * I$, is applied to the
integer. The top and bottom row of each expression evaluate to the
same integer. If we consider only the first column, the top tree has 4
unique subtrees and the bottom has 7. But each column adds 3 unique
subtrees to the top row and only 2 to the bottom. So by the fifth
column the top row has more unique subtrees than the bottom row. The
increased compressibility of the bottom row is due to the fact that
the squaring operation has been factored out as an abstraction. This
is an example of how this compressibility metric favors
abstraction when multiple expressions are being considered at once.

@@

Given a set of chosen solution expressions, which we will call the
\emph{solution set}, want to re-estimate our stochastic grammar. Our
inspiration for doing this is the Neville-Manning algorithm for
grammar-based compression~\cite{nevill1997identifying}. The idea of
that compression algorithm is to compress any string by creating a
grammar for that string such that a) no two symbols appear more than
once in the grammar and b) every rule is used more than once. From a
set of expressions, we generate a grammar according to these criteria.

@@

\setcounter{subfigure}{0} 
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\linewidth]{figures/symreg3frontiers.pdf}
\vspace{-.2cm}
\caption{Learning curves as a function of frontier size.  As frontier
  size is increased, curves plateau closer to $100\%$ performance. A
  baseline search over $150000$ expressions only hits $3\%$ of the
  tasks. \label{fig:symregLearning}}
\vspace{-.3cm}
\end{figure}
\begin{figure}[ht]
\includegraphics[width=1.07\linewidth]{figures/symRegDiffCurricula.pdf}
\caption{How do different task sets affect learning curves? Learning
  curves at frontier size of $10000$ for different task
  sets.} 
\label{fig:symregCurricula}

@@

In Figure~\ref{fig:symregLearning}, we show performance results as we
vary the frontier size.Changing the frontier size changes the number
of tasks the algorithm identifies correctly over the course of 15
algorithm iterations (all experiments in this work were run for 15
iterations). As a baseline, we enumerated $10000*15=150000$
expressions from the initial grammar. This is the total number of
expressions that a run of the algorithm sees if it has a frontier size
of $10000$ and runs for $15$ iterations. Thus, if the benefit accruing
to the runs with larger frontier sizes is simply due to an increase in
the number of expressions seen, rather than increased learning, we
should see similarly good performance from this baseline run. In fact,
however, the baseline run only hits $27$ of the $1000$ tasks ($3\%$),
whereas our algorithm nears $100\%$ for a frontier of $10000$.

What does the E.C. algorithm learn in this task?  Inspecting the top
weighted primitives of the final grammars, we find many incrementers
(e.g. (+1), (+2), etc.), several versions of expressions that
translate to functions like $x*f(x)$ and $f(x*f(x))$, and combinations
of these, like $x*(x+3)+3$ and $x*(x+2)$. Informally, we see
expressions that apply functions to themselves, building up complex
functions with relatively few unique primitives.

To what degree is ``conceptual bootstrapping'' responsible for the
improved performance. That is, to what extent does the presence of
simple problems account for the ability to learn complex functions?
One hypothesis is that to succeed on a set of tasks, the set must
contain a ``learnable'' curriculum, a set of tasks that serve as
stepping stones from an impoverished representation to a rich one. We
can test this hypothesis by varying the set of tasks to which the E.C.
algorithm is exposed. If it is true, then we should see a nonlinear
response to reducing the number of easy tasks, as the curriculum
changes from being learnable to being unlearnable.

@@

\begin{figure}[t]
\includegraphics{./figures/tablealonedoc.pdf}
%% \vspace{-.5cm}
\caption{Some learned primitives from the circuit based Boolean
  function learning experiment. We highlight compression using $NOT$,
  $E_1$ and $E_2$. \,(a)The three elementary gates used to generate
  the task set. (b) Two higher-order primitives in the induced
  grammar. Note how $E_1$ can be used to define FALSE by taking the
  NOT function as an argument. (c) Additional common Boolean functions
  found in the learned grammar.  \label{table:booleanCircuits}}
%%\vspace{-.4cm}
\end{figure}

As another demonstration of our approach, we investigate the
E.C. algorithm's ability to learn Boolean functions using only the logical
primitive NAND and the basic combinators. It is well known that a
boolean circuit can be constructed for any boolean function given only
the NAND gate. Here, the basic combinators take the place of the
wiring normally found in a circuit.

@@

We constructed two task sets. The first of these was constructed
explicitly to contain familiar modular structure. In this set of
tasks, we sampled $1000$ Boolean circuits using AND, OR, and NOT
gates. To accomplish this, we first sampled either 1, 2, 3 inputs,
then between 1 and 5 gates, randomly wiring the inputs of each new
gate to the output of one of the existing gates in the circuit. We
continued this sampling procedure until we had 1000 ``connected''
circuits, i.e., circuits all of whose outputs are wired to an input
except for the last one (the output gate). This resulted in $1000$
tasks consisting of $82$ unique Boolean functions, with a distribution
of tasks as shown in Figure~\ref{fig:booldistr}~(a). In
Figure~\ref{fig:boolLearningCurvesCircuits}, we present learning
curves for frontier sizes 100, 500 and 10000, using the same procedure
as in Section~\ref{sec:symreg}. In Figure~\ref{fig:booldistr}, we show
that the distribution over Boolean functions enumerated from the
grammar over expressions is much more similar to the true function
distribution after learning than before.

\begin{figure}
%%\vspace{-.3cm}
\includegraphics[width=\linewidth]{figures/circuitDistrComp.pdf}
%%\vspace{-.6cm}
\caption{Distribution of boolean functions among (a) the set of 1000
  circuits randomly constructed from AND, OR and NOT gates (see text);
  b) the first 1000 expressions from the grammar over expressions
  \emph{before} learning; c) the first 1000 expressions from the
  grammar over expressions \emph{after} learning. Gray levels indicate
  percent of instances (circuits/expressions) which evaluate to that
  function.\label{fig:booldistr}}
%%\vspace{-.4cm}
\end{figure}

This problem is particularly suitable for inspecting the constituents
of the induced grammar $\mathcal{G}$. We might hope that our algorithm
recovers the constituent logic gates that were used to build up the
tasks. Table~\ref{table:booleanCircuits} shows a few of the top ranked
expressions in the library. These include the basic logic gates; we
also include two higher-order expression $E_1$ and $E_2$, to stress
that the representation the E.C. algorithm is using allows it to build
concepts that are more expressive that just sub-circuits.

In a second experiment, we include all $272$ Boolean truth tables with
three or fewer inputs as tasks. This is a more difficult
problem. There are two kinds of learning that the E.C. algorithm might
accomplish: first, many expressions in $\mathcal{L}$ map to a single
Boolean function, so it needs to learn primitives that allow it to
span many different functions instead of generating very simple
functions redundantly. The second kind of learning involves the
distribution of the functions themselves. In the circuit based Boolean
experiment, that structure is apparent in the distribution in
Figure~\ref{fig:booldistr}. In this second experiment, we remove the
second kind of structure. In
Figure~\ref{fig:boolLearningCurvesUniform}, we show 10 runs of the
E.C.  algorithm on the second experiment with a frontier size of
$2000$: note how there are several local minima that the majority of
the runs get stuck in with performance around $50\%$, but several of
the runs seem to take different trajectories to more successful
representations.

\vspace{-.3cm}
\begin{figure}[h]
\begin{subfigure}{.1\linewidth}
\caption{\label{fig:boolLearningCurvesCircuits} }
\end{subfigure}
\begin{subfigure}[Before]{0.9\linewidth}
\includegraphics[width=1.0\linewidth]{figures/booleancircuits3frontiers.pdf}
\end{subfigure}

\vspace{-.32cm}
\begin{subfigure}{.1\linewidth}
\caption{\label{fig:boolLearningCurvesUniform}}
\end{subfigure}
\begin{subfigure}[After]{.9\linewidth}
\includegraphics[width=1.0\linewidth]{figures/booleanfunctionsUniform.pdf}

@@

Our specific implementation of this approach involved a number of
choices which should be explored further in future work.  Our claim
here is not that the E.C. algorithm is optimal, but that it captures key
features of how people solve the problem of learning new systems of
concepts in acquiring deep domain expertise.  Much like a human
researcher investigating a new area of study, our algorithm tries
different models, watches them fail or succeed, and abstracts out the
model elements that capture relevant differences.  It recombines and
reuses successful patterns of reasoning that span many problems in a
domain.
